Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-13 paper/sheaf_lics.tex
explicitize proof of (o(a=b)=(mu a = mu b))

\newtheorem{thm}{Theorem}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{defi}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma}

@@

Our work is mainly based on the stratification by $\Type_n$~:
\begin{defi}
  $\IsType n$ is defined by induction on $n\geqslant -2$~:
  \begin{itemize}
  \item $\IsType {(-2)} X$ if $X$ is a contractible type, {\em ie} $X$
    is pointed by $c:X$, and every other point in $X$ is connected to $c$.
  \item $\IsType {(n+1)} X = \prod_{x,y:X} \IsType n (x=y).$
  \end{itemize}
  Then, $\Type_n = \sum_{X:\Type} \IsType n X$.
\end{defi}
For a type $X$, to be in $\Type_n$ means that path spaces of $X$ are
trivial after $n+1$ iteration.
We define some shortcuts
\begin{itemize}
\item $\IsContr := \IsType {(-2)}$ and $\Contr := \Type_{-2}$
\item $\IsHProp := \IsType {(-1)}$ and $\HProp := \Type_{-1}$
\end{itemize}

@@

\subsection{Generalized pullbacks}
\label{sec:gener-pb}

@@

The Giraud axiom asserts that, if $f:X\to Y$ is a surjection, then the
colimit of $C(f)$ is $Y$. Viewing homotopy type theory as a
type-theoristic version of higher topos theory, \cite{lurie} suggests that this
axiom can be proved, although we admit it in this paper.

@@

\begin{defi}
  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $A$ of $E$
  classified by $\chi$ is {\em dense} in $E$ if its $\modal$-closure
  is $E$ seen as a subobject of $E$, {\em ie}
  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal (\chi~e)).$$ 
  Moreover, for all $x:A$ we need the following coherence diagram to
  commute
  $$\xymatrix{
    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_\iota & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
    \sum_{e':E} x = e' \ar@{=}[r] & \modal (\chi~x)
  }$$
  where $\iota: x \mapsto ({x_1}_1 ; x_2)$.
\end{defi}

@@

We mimic the construction in~\cite{maclanemoerdijk}

\subsection{For h-propositions}
\label{ssec:h-propositions}

@@

@@

\begin{proof}
  As $\mu_T$ is an embedding, we know by Giraud axiom that $\square T$ is the colimit
  of $C(\mu_T)$. If we can show that $C(\mu_T) = C(\id)$, the the
  result will follow.

  Let $k:\nat$, let's show that 
  \begin{align*}
    &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
      \land \mu_T t_{k-1} \mu_T t_k)\\
    &= \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
      \land t_{k-1} = t_k)
  \end{align*}
By induction on $k$, and the preservation of products by $\modal$, it
suffices to show that for all $a,b:T$, $\modal (a=b) = (\mu_T a =
\mu_T b)$, which is true by univalence.
\end{proof}

@@

\bibitem[HTT]{lurie}
Jacob Lurie
\newblock{\em Higher Topos Theory}

\bibitem[lumsdaine]{lumsdaine}
Lumsdaine
\newblock{Diagrams and limits}

>>>>

\newtheorem{thm}{Theorem}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{defi}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{ax}[thm]{Axiom}

@@

Our work is mainly based on the stratification by $\Type_n$~:
\begin{defi}
  $\IsType n$ is defined by induction on $n\geqslant -2$~:
  \begin{itemize}
  \item $\IsType {(-2)} X$ if $X$ is a contractible type, \ie $X$
    is pointed by $c:X$, and every other point in $X$ is connected to $c$.
  \item $\IsType {(n+1)} X = \prod_{x,y:X} \IsType n (x=y).$
  \end{itemize}
  Then, $\Type_n = \sum_{X:\Type} \IsType n X$.
\end{defi}
For a type $X$, to be in $\Type_n$ means that path spaces of $X$ are
trivial after $n+1$ iteration.
We define some shortcuts
\begin{itemize}
\item $\IsContr := \IsType {(-2)}$ and $\Contr := \Type_{-2}$
\item $\IsHProp := \IsType {(-1)}$ and $\HProp := \Type_{-1}$
\end{itemize}

@@

\subsection{Generalized pullbacks}
\label{sec:gener-pb}
\kq{Do we keep this name ?}

@@

\begin{ax}[Giraud]
  If $f:X\to Y$ is a surjection, then the
colimit of $C(f)$ is $Y$.
\end{ax}

Viewing homotopy type theory as a
type-theoristic version of higher topos theory, \cite{lurie} suggests that this
axiom can be proved, although we admit it in this paper.

Formalization of colimits is based on the formalization of limits in~\cite{lumsdaine}.

@@

\begin{defi}
  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $A$ of $E$
  classified by $\chi$ is {\em dense} in $E$ if its $\modal$-closure
  is $E$ seen as a subobject of $E$, \ie
  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal (\chi~e)).$$ 
  Moreover, for all $x:A$ we need the following coherence diagram to
  commute
  $$\xymatrix{
    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_\iota & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
    \sum_{e':E} x = e' \ar@{=}[r] & \modal (\chi~x)
  }$$
  where $\iota: x \mapsto ({x_1}_1 ; x_2)$.
\end{defi}

@@

We mimic the construction in~\cite{maclanemoerdijk}. 
From any left exact modality on $\HProp$, we extend the new principles
it gives to every $\Type_n,\,n\geqslant 0$.
Actually, if $n_0$ is a fixed truncation index, and $\modal$ a left
exact modality on $\Type_{n_0}$, then we can in the same
way extend the properties of the modality to any $\Type_n$, for
$n > n_0$.

Once the first modality is defined, the extension to all $\Type_n$ is
automatic. It means that the new principles we want to add in the new
theory must be introduced ine the first modality.

The sheafification process works exactly that way
in~\cite{maclanemoerdijk}~: from a left exact modality on the internal
logic (a Lawvere-Tierney topology), we define a new left exact
modality on the whole topos. We view that as the induction step
between $\HProp$ and $\Type_0$, and we want to extend it.

\kq{Bof ; maybe rewrite this}

As in most of sheafification process, it will be done in
two steps~:
\begin{enumerate}[(i)]
\item {\em separation~:} In this step, we {\em remove} the lack of
  sheafness from the type we are considering.
\item {\em completion~:} In this step, we add what lacks now to the
  separated type to be a sheaf.
\end{enumerate}
\subsection{For h-propositions}
\label{ssec:h-propositions}

@@

The $\lnot\lnot$-modality allows to work with a propositional law of
excluded middle.

@@

\begin{proof}
  As $\mu_T$ is an surjection, we know by Giraud axiom that $\square T$ is the colimit
  of $C(\mu_T)$. If we can show that $C(\mu_T) = C(\id)$, the the
  result will follow.

  Let $k:\nat$, let's show that 
  \begin{align*}
    &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
      \land \mu_T t_{k-1} \mu_T t_k)\\
    &= \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
      \land t_{k-1} = t_k)
  \end{align*}
By induction on $k$, and the preservation of products by $\modal$, it
suffices to show that for all $a,b:T$, $\modal (a=b) = (\mu_T a =
\mu_T b)$. By univalence, we want arrows in both ways, forming an
equivalence.
\begin{itemize}
\item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
  first components yields $q : \prod_{t:T} \modal(a=t) = \modal (b=t)
  $.
  Taking for example $t=b$, we deduce $\modal (a=b) = \modal(a=a)$,
  and the latter is inhabited by $\eta_{a=a} 1$.
\item Suppose now $p : \modal(a=b)$. Let $\iota$ be the first
  projection from $\square T \to (T \to \Type_n^\modal)$. $\iota$ is
  an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
  (\mu_T b)$, \ie $\prod_{t:T}\modal (a=t) = \modal (b=t)$. The latter
  remains true by univalence.
\end{itemize}
The fact that these two form an equivalence is technical, we refer to
the formalization for an explicit proof.
\end{proof}

@@

\bibitem[HTT]{lurie}
Jacob Lurie.
\newblock{\em Higher Topos Theory}.

\bibitem[AKL]{lumsdaine}
Peter LeFanu Lumsdaine, Jeremy Avigad and Chris Kapulkin.
\newblock{\em Homotopy limits in type theory}.
